\documentclass{article}\usepackage{amsmath}\begin{document}\title{Simplex Solver}\maketitle\begin{flushleft}\textbf{Problem}\end{flushleft}\begin{flushleft}Given the following linear system and objective function, find the optimal solution.\end{flushleft}\begin{equation*}\max{ x_1 + x_2 + x_3 + x_4 } \\ \end{equation*}\[\left\{\begin{array}{c} y_1 + 2y_2 - y_3 - y_4 \geq 1 \\  y_1  -5y_2 + 2y_3 + 3y_4 \geq 1 \\ \end{array}\right.\]\begin{flushleft}\textbf{Solution}\end{flushleft}\begin{flushleft}Add slack variables to turn all inequalities to equalities.\end{flushleft}\[\left\{\begin{array}{c} y_1 + y_2 + s_1 = 1 \\  2y_1  -5y_2 + s_2 = 1 \\ - y_1 + 2y_2 + s_3 = 1 \\ - y_1 + 3y_2 + s_4 = 1 \\ \end{array}\right.\]\begin{flushleft}Create the initial tableau of the new linear system.\end{flushleft}\begin{equation*}\begin{bmatrix}\begin{array}{cccccc|c}y_1 &y_2 &s_1 &s_2 &s_3 &s_4 &b \\ \hline1 & 1 & 1 & 0 & 0 & 0 & 1 \\2 & -5 & 0 & 1 & 0 & 0 & 1 \\-1 & 2 & 0 & 0 & 1 & 0 & 1 \\-1 & 3 & 0 & 0 & 0 & 1 & 1 \\ \hline-1 & -1 & 0 & 0 & 0 & 0 & 0 \\\end{array}\end{bmatrix}\begin{array}{c}\\s_1 \\s_2 \\s_3 \\s_4 \\\\\end{array}\end{equation*}\begin{flushleft}There are negative elements in the bottom row, so the current solution is not optimal. Thus, pivot to improve the current solution. The entering variable is $y_1$ and the departing variable is $s_2$.\end{flushleft}\begin{flushleft}Perform elementary row operations until the pivot element is 1 and all other elements in the entering column are 0.\end{flushleft}\begin{equation*}\begin{bmatrix}\begin{array}{cccccc|c}y_1 &y_2 &s_1 &s_2 &s_3 &s_4 &b \\ \hline0 & 7/2 & 1 & -1/2 & 0 & 0 & 1/2 \\1 & -5/2 & 0 & 1/2 & 0 & 0 & 1/2 \\0 & -1/2 & 0 & 1/2 & 1 & 0 & 3/2 \\0 & 1/2 & 0 & 1/2 & 0 & 1 & 3/2 \\ \hline0 & -7/2 & 0 & 1/2 & 0 & 0 & 1/2 \\\end{array}\end{bmatrix}\begin{array}{c}\\s_1 \\y_1 \\s_3 \\s_4 \\\\\end{array}\end{equation*}\begin{flushleft}There are negative elements in the bottom row, so the current solution is not optimal. Thus, pivot to improve the current solution. The entering variable is $y_2$ and the departing variable is $s_1$.\end{flushleft}\begin{flushleft}Perform elementary row operations until the pivot element is 1 and all other elements in the entering column are 0.\end{flushleft}\begin{equation*}\begin{bmatrix}\begin{array}{cccccc|c}y_1 &y_2 &s_1 &s_2 &s_3 &s_4 &b \\ \hline0 & 1 & 2/7 & -1/7 & 0 & 0 & 1/7 \\1 & 0 & 5/7 & 1/7 & 0 & 0 & 6/7 \\0 & 0 & 1/7 & 3/7 & 1 & 0 & 11/7 \\0 & 0 & -1/7 & 4/7 & 0 & 1 & 10/7 \\ \hline0 & 0 & 1 & 0 & 0 & 0 & 1 \\\end{array}\end{bmatrix}\begin{array}{c}\\y_2 \\y_1 \\s_3 \\s_4 \\\\\end{array}\end{equation*}\begin{flushleft}There are no negative elements in the bottom row, so we know the solution is optimal. Thus, the solution is: \end{flushleft}\begin{equation*}s_1 = 0, s_2 = 0, s_3 = \frac{11}{7}, s_4 = \frac{10}{7}, x_1 = 1, x_2 = 0, x_3 = 0, x_4 = 0, y_1 = \frac{6}{7}, y_2 = \frac{1}{7}, z = 1\end{equation*}\end{document}